Discriminative Sensing Based on Signal Processing for Information Security Analysis

### 2017-11-14 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
improving the approach and experiment description

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detecton from imbalanced data, in order to apply the tensor deconposition for DL methods to evidentiate the discriminative sensing of a fraud detection dataset.

This chapter is organized as follows. Section \ref{sec:4_motivation} presents the motivation for the proposed work. In Section \ref{sec:4_relatedworks}, related works are discussed. Section \ref{sec:4_datamodel} presents the data model and the evaluated datasets. Section \ref{sec:4_proposal} describes the proposed approach for for fraud detecton from mobile money transacion. Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.

@@

MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}. MOD defines that if we know $S$ in equation \ref{eq:4_eq01}, $A$ can be estimated solving a linear LS problem. Additionally, knowing $A$ we obtain $S$ by solving a sparse recovery problem. MOD iterates between the two steps, starting with an initial guess of $A$, denoted as $\hat{A}$, which can be estimated, generated by previous dictionary learning or generated analitically. Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $A$ and $S$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.

K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning. After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $A$ using SVD to minimize the approximation error. The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights. In practice, dictionaries learned with K-SVD
have shown excellent performance in image denoising.

@@

Dimensionality reduction can first be achieved by selecting a subset of functions from a large, fixed dictionary that is used for the analysis of particular signals. These functions then determine a subspace of reduced dimension, where classification can be performed by computing the nearest neighbor points among the projected data. A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace. This kind of approach typically tries to maximize the variance between the active atoms from $A$ that represent signals in different classes. The reduced subspace used for classification is finally formed by the subset of atoms in $A$ whose corresponding coefficients in $S$ are nonzero. The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.

An important advantage of redundant dictionaries for classification is that signal analysis can be performed with functions that are likely to match the data characteristics in different classes of signals. Similarly to data approximation problems, data analysis applications can further benefit from dictionary learning methods. The previous section describes subspace selection methods from predefined dictionaries. DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes. The atoms in a dictionary $A$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals. They naturally permit to classify the data into the corresponding linear subspace. However, there is no guarantee that the subspace built on a learned dictionary $A$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.

@@

\subsection{Dataset}
\label{sec:4_dataset}

We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal, this dataset is synthetically generated by  PaySim \cite{lopez2016paysim} , which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem. PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behaviour to later evaluate the performance of fraud detection methods. PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country. The original logs were provided by a multinational company

The selected dataset is composed by 9 features (variables) and 6.362.620 tramsactions (observations), it is also provided one lable that classifies each transaction as fraudulent or not. As the majority dataset of anomaly or fraud detection, the data is imbalanced, where the number of fraudlent transactions is ZZ against of NN normal transactions.

Imbalance data where zero is predominante

Data divided into learning and test, with isfraud labels with 1 or 0

\subsection{Tensor-based Model}
\label{sec:4_tensor_model}

The Kronecker model (\ref{eq:4_eq02}) can be rewritten in an equivalent tensor form. Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:4_eq02}) into a "Tucker-2" decomposition, as

\begin{equation}\label{eq:4_eq04}
\boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
\end{equation}

where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.

We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained. In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons. We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.

The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.

\section{Proposed Approach}
\label{sec:4_proposal}

...

@@

rocauc is usually used for classification evaluation, but It os not the best option for imbalance data, and Also because It does not consider the unclassified cases, such as false negatives

Receiver Operating Characteristic Curve (ROC curve) http://www.chioka.in/differences-between-roc-auc-and-pr-auc/

A ROC curve is plotting True Positive Rate (TPR) against False Positive Rate (FPR).

TPR is defined as:

tpr-formula

FPR is defined as:

fpr-formula

where TP = true positive, TN = true negative, FP = false positive, FN = false negative.

A typical ROC curve looks like this, which shows two ROC curves for Algorithm 1 and Algorithm 2.

sample-ROC-curve

The goal is to have a model be at the upper left corner, which is basically getting no false positives – a perfect classifier.

The receiver operating characteristic area under curve (ROC AUC) is just the area under the ROC curve. The higher it is, the better the model is.

Precision Recall Curve (PR Curve)
A PR curve is plotting Precision against Recall.

Precision is defined as:

precision-formula

Recall is defined as:

recall-formulaA typical PR curve looks like this, which shows two PR curves for Algorithm 1 and Algorithm 2.

sample-PR-curve

The goal is to have a model be at the upper right corner, which is basically getting only the true positives with no false positives and no false negatives – a perfect classifier.

The precision recall area under curve (PR AUC) is just the area under the PR curve. The higher it is, the better the model is.

Differences between the ROC AUC and PR AUC
Randy has given a great explanation here, plus a little of my understanding.

Since PR does not account for true negatives (as TN is not a component of either Precision or Recall), or there are many more negatives than positives (a characteristic of class imbalance problem), use PR. If not, use ROC.

Due to the inherent complex characteristics of imbalanced data sets, learning from such data requires new understandings, principles, algorithms, and tools to transform vast amounts of raw data efficiently into information and knowledge representation.

We adopt precision recall curve and compare with logistic regression based on original and pca data

Classification based on reconstruction or other value, such as similarity analysis

Evaluate what metric is more discriminative, mse, tsnr or other

@@

>>>>

Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor deconposition for DL methods to evidentiate the discriminative sensing of a fraud detection dataset.

This chapter is organized as follows. Section \ref{sec:4_motivation} presents one experiment and results that motivate the proposed work. In Section \ref{sec:4_relatedworks}, related works are discussed. Section \ref{sec:4_datamodel} presents the data model and the evaluated datasets. Section \ref{sec:4_proposal} describes the proposed approach for for fraud detecton from mobile money transacion. Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.

@@

MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}. MOD defines that if we know $\boldsymbol{S}$ in equation \ref{eq:4_eq01}, $\boldsymbol{A}$ can be estimated solving a linear LS problem. Additionally, knowing $\boldsymbol{A}$ we obtain $\boldsymbol{S}$ by solving a sparse recovery problem. MOD iterates between the two steps, starting with an initial guess of $\boldsymbol{A}$, denoted as $\hat{\boldsymbol{A}}$, which can be estimated, generated by previous dictionary learning or generated analitically. Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\boldsymbol{A}$ and $\boldsymbol{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.

K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning. After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $\boldsymbol{A}$ using SVD to minimize the approximation error. The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights. In practice, dictionaries learned with K-SVD
have shown excellent performance in image denoising.

@@

Dimensionality reduction can first be achieved by selecting a subset of functions from a large, fixed dictionary that is used for the analysis of particular signals. These functions then determine a subspace of reduced dimension, where classification can be performed by computing the nearest neighbor points among the projected data. A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace. This kind of approach typically tries to maximize the variance between the active atoms from $\boldsymbol{A}$ that represent signals in different classes. The reduced subspace used for classification is finally formed by the subset of atoms in $\boldsymbol{A}$ whose corresponding coefficients in $\boldsymbol{S}$ are nonzero. The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.

An important advantage of redundant dictionaries for classification is that signal analysis can be performed with functions that are likely to match the data characteristics in different classes of signals. Similarly to data approximation problems, data analysis applications can further benefit from dictionary learning methods. The previous section describes subspace selection methods from predefined dictionaries. DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes. The atoms in a dictionary $\boldsymbol{A}$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals. They naturally permit to classify the data into the corresponding linear subspace. However, there is no guarantee that the subspace built on a learned dictionary $\boldsymbol{A}$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.

@@

% \subsection{Dataset}
% \label{sec:4_dataset}

We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal. This dataset is synthetically generated by  PaySim \cite{lopez2016paysim}, which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem. PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behaviour to later evaluate the performance of fraud detection methods. PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country. The original logs were provided by a multinational company

The selected dataset is composed by 9 features (variables) and 2.770.408 transactions (observations). It is also provided one lable that classifies each transaction as fraudulent or not. As the majority dataset of anomaly or fraud detection, the data is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.196 of normal transactions. Therefore, the fraud class represents 0,30 \% of the entire datased.

For evaluation of the proposed classification approach we adopt a data splitting into training and test data, where the training data has 70 \% of the whole dataset. We denote the traning data as $X_r$ and the test data as $X_e$. The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $X_r^o$, the fraud train data is denoted as $X_r^f$. The same notation of fraud, normal, train and test is applied to qualify $A$ and $S$.

% \subsection{Tensor-based Model}
% \label{sec:4_tensor_model}

% The Kronecker model (\ref{eq:4_eq02}) can be rewritten in an equivalent tensor form. Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:4_eq02}) into a "Tucker-2" decomposition, as

% \begin{equation}\label{eq:4_eq04}
% \boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
% \end{equation}

% where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.

% We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained. In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons. We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.

% The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.

\section{Tensor-based Sparce Representation Classification}
\label{sec:4_proposal}

We propose a Tensor-based Sparce Representation Classification approach to use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.

Our proposal requires supervision, where labels of training data are used to classify each transaction as fraud or normal. The first step is to split the training dataset into classes, as fraud or normal, and create respectively the data matrices $\boldsymbol{X}_r^f$ and $\boldsymbol{X}_r^o$. For each data matrix we compute the data normalization and subsequently the initial dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$. Afterward, we apply the dictionaries and the data matrices to the selected DL algorithms, in order to learn the best dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ that respectivelly represents fraud and normal classes as basis vectors.

With the trainned dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ and the test data $\boldsymbol{X}_e$ it is possible to process each transacion of $\boldsymbol{X}_e$, denoted as $\boldsymbol{x}_e^t$ where $t$ represents the transaction index, in order to compute its sparse coding from fraud and normal dictionaries, generating respectively $\boldsymbol{S}_e^f$ and $\boldsymbol{S}_e^o$. Subsequently, we compute the signal rescontruction as $\boldsymbol{x}_e^f = \boldsymbol{A}_r^f \cdot \boldsymbol{S}_e^f$ and $\boldsymbol{x}_e^o = \boldsymbol{A}_r^o \cdot \boldsymbol{S}_e^o$.

Finally, we compare $\boldsymbol{x}_e^f$ and $\boldsymbol{x}_e^o$ to $\boldsymbol{x}_e^t$ in order to evaluate the reconstruction error to classify the transaction as fraud or normal according to the minimum reconstruction error from $\boldsymbol{x}_e^f$ and $\boldsymbol{x}_e^o$, respectivelly. We adopt the cosine similarity to compare the reconstructed signal to the original one.

@@

For data analysis of imbalanced classes it is usually applied a data undersample method, where an under sized sample is selected from the original datased, with similar amount of occurrences for each class of the data. This method is efficient for algorithm trainning, but it faces generalization problems when tries to classify new data. Therefore, we propose to evaluate the our proposal for under sampled data and for the complete dataset.

The Receiver Operating Characteristics (ROC) curve makes use of the proportion of two metrics: true positives rate (TPR) and false positives rate (FPR), which are defined as:

\begin{equation}\label{eq:4_eq04}
	TPR = \frac{TP}{Pc},
\end{equation}

\begin{equation}\label{eq:4_eq05}
	FPR = \frac{FP}{Nc},
\end{equation}

where $TP$ denotes true positive, $Pc$ is the total positive values, $FP$ denotes the false positive and $Nc$ the total negative values.

ROC curves are commonly used to present results for binary decision problems in machine learning. However, when dealing with highly skewed datasets, Precision-Recall (PR) curves give a more informative picture of an algorithm’s performance \cite{davis2006relationship, he2009learning}. The PR curve is defined by plotting Precision Rate (PR) over the Recall Rate (RR), which are defined as:

\begin{equation}\label{eq:4_eq06}
	PR = \frac{TP}{TP + FP},
\end{equation}

\begin{equation}\label{eq:4_eq07}
	RR = \frac{TP}{TP + FN}.
\end{equation}

PR curves exhibit a strong correspondence to ROC curves: A curve dominates in ROC space if and only if it dominates in PR space. However, an algorithm that optimizes the AUC in the ROC space is not guaranteed to optimize the AUC in PR space. Considering that our data is very imbalanced, with many more negatives than positives values, we adopt the averate precision recall for the classification evaluation

Logistic regression is the most common method used to model binary response data \cite{hilbe2011logistic}, it is useful for situations in where it is necessary to predict the presence or absence of a characteristic, with binary classification, such as in anomaly or fraud detection \cite{bhattacharyya2011data}. Therefore, we compare the results of out proposed approach with the logistic regression algorithm applied for fraud detection from the adopted datase.

\section{Results}
\label{sec:4_results}

% \section{Discussion}
% \label{sec:4_Discussion}

%problem of imbalanced data

@@

incorporate discriminative funcions

evaluate other metrics beyond reconstruction, using techniques such as reconstruction with dictionary concatenation
